---
title: "BGD modeling simulations"
author: "Austin Schumacher"
date: "`r Sys.Date()`"
output: 
    pdf_document:
        includes: 
            in_header: "/Users/austin/Desktop/work/presentation_publication_resources/mycommands.tex"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = FALSE)
rm(list = ls())

## Libraries ####
library(SUMMER)
# help(package = "SUMMER", help_type = "html")
library(classInt)
library(RColorBrewer)
library(dplyr)
library(tidyr)
library(rgdal)
library(scales)
library(survey)
library(tidyverse)
library(spdep)
library(INLA)
library(knitr)

set.seed(858)
```

# Continuous outcome

## IID (nonspatial)

### No shared components

We will start by simulating continuous outcomes for 2 "causes" with the following IID model:

\eq{
    y_{ic} &= \beta_c + \gamma_{i} + \epsilon_{ic} \\
    \epsilon_{ic} &\overset{iid}{\sim} N(0, \sigma_{\epsilon}^2) \\
    \gamma_{i} &\overset{iid}{\sim} N(0, \sigma_{\gamma}^2)
}

First, we simulate data for $i = 1, \dots, 8$ regions and $c = 1, \dots, 2$ causes, with

- $\bm\beta = (1,2)$
- $\sigma_{\epsilon}^2 = 0.1$
- $\sigma_{\gamma}^2 = 0.5$

We fit a correctly specified model with

- flat priors for each $\beta_c$
- Gamma(1, 0.00005) prior on $1 / \sigma^2_{\epsilon}$
- PC prior for $\sigma_{\gamma}$ such that there is a 1\% chance $\sigma_{\gamma} > 1$

```{r}
# parameters
nreg <- 8
ncause <- 2
beta <- c(1,2)
gamma <- rnorm(nreg, 0, 0.5)
gamma1 <- rnorm(nreg, 0, 0.5)
gamma2 <- rnorm(nreg, 0, 0.5)
sigma <- 0.1

# simulate data
# nsims <- 10
# for (i in 1:nsims) {
#     
# }
y1 <- rnorm(nreg, beta[1] + gamma, sigma)
y2 <- rnorm(nreg, beta[2] + gamma, sigma)
# y1 <- rnorm(nreg, beta[1] + gamma1, sigma^2)
# y2 <- rnorm(nreg, beta[2] + gamma2, sigma^2)
y <- c(y1, y2)
cause <- rep(c(1,2), each = nreg)
region <-  rep(1:nreg, 2)
region1 <- c(1:nreg, rep(NA, nreg))
region2 <- c(rep(NA, nreg), 1:nreg)

# priors
fe.prec <- list(prec.intercept = 0,
                prec = 0)
iid_prec_prior <- list(prec = list(prior = "pc.prec", param = c(1, 0.01)))

# model formula
m.form <- y ~ -1 + factor(cause) + f(region, model = "iid", hyper = iid_prec_prior)
# m.form <- y ~ -1 + factor(cause) + 
#     f(region1, model = "iid", hyper = iid_prec_prior) +
#     f(region2, model = "iid", hyper = iid_prec_prior)

# fit model
mod1 <- inla(m.form, 
             data = data.frame(y = y, cause = cause, region = region, 
                               region1 = region1, region2 = region2),
             family = "gaussian",
             control.fixed = fe.prec,
             control.predictor=list(compute=TRUE),
             control.compute=list(config = TRUE))
fe.res <- mod1$summary.fixed
re.res <- mod1$summary.random
hyperpar.res <- mod1$summary.hyperpar
fitted.res <- mod1$summary.fitted.values

res.table <- data.frame(parameter = c("$\\beta_1$", "$\\beta_2$", 
                         "$\\sigma_{\\epsilon}$", "$\\sigma_{\\gamma}$"),
                        med = c(fe.res$`0.5quant`, 1/(hyperpar.res$`0.5quant`)^0.5),
                        L95 = c(fe.res$`0.025quant`, 1/(hyperpar.res$`0.975quant`)^0.5),
                        U95 = c(fe.res$`0.975quant`, 1/(hyperpar.res$`0.025quant`)^0.5))
kable(res.table)

plot(fitted.res$`0.5quant` ~ y, main = "Comparison of fitted values with 95% UI",
     xlab = "observed", ylab = "fitted",
     ylim = range(c(fitted.res$`0.025quant`, fitted.res$`0.975quant`)))
segments(y, fitted.res$`0.025quant`, y, fitted.res$`0.975quant`)
abline(0,1, col = "red")

plot(re.res$region$`0.5quant` ~ gamma, main = "Comparison of REs with 95% UI",
     xlab = "observed", ylab = "fitted", 
     ylim = range(c(re.res$region$`0.025quant`, re.res$region$`0.975quant`)))
segments(gamma, re.res$region$`0.025quant`, gamma, re.res$region$`0.975quant`)
abline(0,1, col = "red")
```

Next, we will move on to modeling 3 "causes":

TODO

### Shared components

We will start by simulating a continuous outcome for 2 "causes" with the following shared component (coregionalization) IID model:

\eq{
    y_{i1} &= \beta_1 + \gamma_{i1} + \epsilon_{i1} \\
    y_{i2} &= \beta_2 + \lambda_1\gamma_{i1} + \gamma_{i2} + \epsilon_{i2} \\
    \epsilon_{ic} &\overset{iid}{\sim} N(0, \sigma_{\epsilon}^2) \\
    \gamma_{i1} &\overset{iid}{\sim} N(0, \sigma_{\gamma}^2) \\
    \gamma_{i2} &\overset{iid}{\sim} N(0, \sigma_{\gamma}^2)
}

First, we simulate data for $i = 1, \dots, n$ regions ($n \in \{ 8, 500\}$) and $c = 1, \dots, 2$ causes, with

- $\bm\beta = (1,2)$
- $\sigma_{\epsilon} = 0.01$
- $\sigma_{\gamma} = 0.5$
- $\lambda_1 = 0.5$ 

We fit a correctly specified model with

- flat priors for each $\beta_c$
- fixed (known) $\sigma_{\epsilon}$ correctly set to 0.01
- PC prior for $\sigma_{\gamma}$ such that there is a 1\% chance $\sigma_{\gamma} > 1$
- N(0,1) prior on $\lambda$ 
  
```{r}
# simulate data
nregs <- c(8, 500)
sigmas <- c(0.01, 0.1)
for (ri in 1:length(nregs)) {
  nreg <- nregs[ri]
  for (si in 1:length(sigmas)) {
    sigma <- sigmas[si]
    
    # parameters
    ncause <- 2
    beta <- c(1,2)
    gamma1 <- rnorm(nreg, 0, 0.5^2)
    gamma2 <- rnorm(nreg, 0, 0.5^2)
    lambda <- 0.5
  
    y1 <- rnorm(nreg, beta[1] + gamma1, sigma)
    y2 <- rnorm(nreg, beta[2] + (gamma1 * lambda) + gamma2, sigma)
    y <- c(y1, y2)
    cause <- rep(c(1,2), each = nreg)
    region <-  rep(1:nreg, 2)
    region1 <- c(1:nreg, rep(NA, nreg))
    region2 <- c(rep(NA, nreg), 1:nreg)
    region12 <- region2
    
    # priors
    fe.prec <- list(prec.intercept = 0,
                    prec = 0)
    iid_prec_prior <- list(prec = list(prior = "pc.prec", param = c(1, 0.01)))
    hyper_lambda <- list(theta = list(prior = 'normal', param = c(0, 1)))
    
    # model formula
    m.form <- y ~ -1 + factor(cause) + 
        f(region1, model = "iid", hyper = iid_prec_prior) +
        f(region2, model = "iid", hyper = iid_prec_prior) +
        f(region12, copy = "region1", fixed = FALSE, hyper = hyper_lambda)
    
    # fit model
    mod1 <- inla(m.form, 
                 data = data.frame(y = y, cause = cause, 
                                   region1 = region1,
                                   region2 = region2,
                                   region12 = region12,
                                   prec_loggaussian = 1/sigma^2),
                 family = "gaussian",
                 control.fixed = fe.prec,
                 control.predictor=list(compute=TRUE),
                 control.compute=list(config = TRUE),
                 control.family = list(hyper = list(prec = list(initial = log(1), fixed=TRUE))),
                 scale = prec_loggaussian)
    
    # results
    fe.res <- mod1$summary.fixed
    re.res <- mod1$summary.random
    hyperpar.res <- mod1$summary.hyperpar
    fitted.res <- mod1$summary.fitted.values
    
    res.table <- data.frame(parameter = c("$\\beta_1$", "$\\beta_2$", 
                             # "$\\sigma_{\\epsilon}$", 
                             "$\\sigma_{\\gamma_1}$", "$\\sigma_{\\gamma_2}$",
                             "$\\lambda$"),
                            med = c(fe.res$`0.5quant`, 1/(hyperpar.res$`0.5quant`[1:2])^0.5, hyperpar.res$`0.5quant`[3]),
                            L95 = c(fe.res$`0.025quant`, 1/(hyperpar.res$`0.975quant`[1:2])^0.5, hyperpar.res$`0.025quant`[3]),
                            U95 = c(fe.res$`0.975quant`, 1/(hyperpar.res$`0.025quant`[1:2]^0.5), hyperpar.res$`0.975quant`[3]))
    kable(res.table, caption = paste0(nreg, " regions; sigma_epsilon = ", sigma))
    
    plot(fitted.res$`0.5quant` ~ y, main = paste0(nreg, " regions: fitted values with 95% UI vs observed"),
         xlab = "observed", ylab = "fitted",
         ylim = range(c(fitted.res$`0.025quant`, fitted.res$`0.975quant`)))
    segments(y, fitted.res$`0.025quant`, y, fitted.res$`0.975quant`)
    abline(0,1, col = "red")
    
    plot(re.res$region1$`0.5quant` ~ gamma1, main = paste0(nreg, " regions: shared RE values with 95% UI vs truth"),
         xlab = "observed", ylab = "fitted", 
         ylim = range(c(re.res$region1$`0.025quant`, re.res$region1$`0.975quant`)))
    segments(gamma1, re.res$region1$`0.025quant`, gamma1, re.res$region1$`0.975quant`)
    abline(0,1, col = "red")
    
    plot(re.res$region2$`0.5quant` ~ gamma2, main = paste0(nreg, " regions: nonshared RE values with 95% UI vs truth"),
         xlab = "observed", ylab = "fitted", 
         ylim = range(c(re.res$region2$`0.025quant`, re.res$region2$`0.975quant`)))
    segments(gamma2, re.res$region2$`0.025quant`, gamma2, re.res$region2$`0.975quant`)
    abline(0,1, col = "red")
  }
}

```

Next, we will model 5 "causes" with a similar model.

## ICAR (spatial)

### No shared components

We will start by simulating continuous outcomes for 2 "causes" with the following BYM spatial model:

\eq{
    y_{ic} &= \beta_c + s_{i} + e_i + \epsilon_{ic} \\
    \epsilon_{ic} &\overset{IID}{\sim} N(0, \sigma^2_{\epsilon})\\
    \bm{s} &\sim ICAR(\sigma_{s}^2) \\
    e_i &\overset{IID}{\sim} N(0, \sigma^2_e) 
}

First, we simulate data for $i = 1, \dots, 8$ regions and $c = 1, \dots, 2$ causes, with

- $\bm\beta = (1,2)$
- $\sigma_{\epsilon}^2 = 0.1$
- $\sigma_{e}^2 = 0.5$
- marginal generalized variance the $s_i$ equal to 1

We fit a correctly specified model with

- flat priors for each $\beta_c$
- Gamma(1, 0.00005) prior on $1 / \sigma^2_{\epsilon}$
- BYM2 specification of the ICAR and IID random effects
  - PC prior on the total standard deviation such that there is a 1\% chance it is bigger than 1
  - PC prior on the proportion of the variance that is spatial such that there is a 50\% chance it is bigger than 0.5

```{r}
## load bgd shape files
poly.adm0 <- readOGR(dsn = "/Users/austin/Dropbox/dissertation_2/survey-csmf/data/shapefiles/bgd_adm_bbs_20201113_SHP",
                     layer = "bgd_admbnda_adm0_bbs_20201113")
poly.adm0$region <- "All"
poly.adm1 <- readOGR(dsn = "/Users/austin/Dropbox/dissertation_2/survey-csmf/data/shapefiles/bgd_adm_bbs_20201113_SHP",
                     layer = "bgd_admbnda_adm1_bbs_20201113")
poly.adm1$region <- tolower(poly.adm1$ADM1_EN)

# get adjacency matrix
admin1.mat.raw <- poly2nb(SpatialPolygons(poly.adm1@polygons))
admin1.mat <- nb2mat(admin1.mat.raw, style = "B")
nb2INLA("bgd.graph", admin1.mat.raw)

# parameters
nreg <- nrow(admin1.mat)
ncause <- 2
beta <- 1:ncause
sigma_epsilon <- 0.1^0.5
sigma_e <- 0.5^.5
sigma_s <- 1^.5

# simulate data
# nsims <- 10
# for (i in 1:nsims) {
#     
# }
e <- rnorm(nreg, 0, sigma_e)
s <- rst(n = 1, type = "s", type.s = "ICAR", Amat = admin1.mat, scale.model = TRUE)
y <- c()
for (b in 1:ncause) {
  y <- c(y, rnorm(nreg, beta[b] + e + s, sigma_epsilon))
}
cause <- rep(1:ncause, each = nreg)
region <-  rep(1:nreg, ncause)

# priors
fe.prec <- list(prec.intercept = 0,
                prec = 0)
bym2_priors <- list(phi=list(prior="pc", param=c(0.5, 0.5), initial=1), 
                    prec=list(prior="pc.prec", param=c(1, 0.01), initial=5))

# model formula
m.form <- y ~ -1 + factor(cause) + f(region, model = "bym2", 
                                     graph = "bgd.graph",
                                     scale.model = TRUE,
                                     constr = TRUE,
                                     hyper = bym2_priors)
# m.form <- y ~ -1 + factor(cause) + 
#     f(region1, model = "iid", hyper = iid_prec_prior) +
#     f(region2, model = "iid", hyper = iid_prec_prior)

# fit model
mod1 <- inla(m.form, 
             data = data.frame(y = y, cause = cause, region = region),
             family = "gaussian",
             control.fixed = fe.prec,
             control.predictor=list(compute=TRUE),
             control.compute=list(config = TRUE))

# results
fe.res <- mod1$summary.fixed
re.res <- mod1$summary.random
hyperpar.res <- mod1$summary.hyperpar
fitted.res <- mod1$summary.fitted.values

res.table <- data.frame(parameter = c("$\\beta_1$", "$\\beta_2$", 
                         "$\\sigma_{\\epsilon}$", "$\\sigma_{T}$", "$\\phi$"),
                        med = c(fe.res$`0.5quant`, 1/(hyperpar.res$`0.5quant`[1:2])^0.5, hyperpar.res$`0.5quant`[3]),
                        L95 = c(fe.res$`0.025quant`, 1/(hyperpar.res$`0.975quant`[1:2])^0.5, hyperpar.res$`0.025quant`[3]),
                        U95 = c(fe.res$`0.975quant`, 1/(hyperpar.res$`0.025quant`[1:2])^0.5, hyperpar.res$`0.975quant`[3]))
kable(res.table)

plot(fitted.res$`0.5quant` ~ y, main = "Comparison of fitted values with 95% UI",
     xlab = "observed", ylab = "fitted",
     ylim = range(c(fitted.res$`0.025quant`, fitted.res$`0.975quant`)))
segments(y, fitted.res$`0.025quant`, y, fitted.res$`0.975quant`)
abline(0,1, col = "red")

# plot(re.res$region$`0.5quant` ~ gamma, main = "Comparison of REs with 95% UI",
#      xlab = "observed", ylab = "fitted", 
#      ylim = range(c(re.res$region$`0.025quant`, re.res$region$`0.975quant`)))
# segments(gamma, re.res$region$`0.025quant`, gamma, re.res$region$`0.975quant`)
# abline(0,1, col = "red")
```

### Shared components

We will start by simulating continuous outcomes for 2 "causes" with the following shared component BYM spatial model:

\eq{
    y_{i1} &= \beta_1 + s_{i1} + e_{i1} + \epsilon_{i1} \\
    y_{i2} &= \beta_2 + \lambda_1 (s_{i1} + e_{i1}) + s_{i2} + e_i + \epsilon_{i2} \\
    \epsilon_{ic} &\overset{IID}{\sim} N(0, \sigma^2_{\epsilon})\\
    \bm{s_c} &\sim ICAR(\sigma_{s}^2) \\
    e_i &\overset{IID}{\sim} N(0, \sigma^2_e) 
}

First, we simulate data for $i = 1, \dots, 8$ regions and $c = 1, \dots, 2$ causes, with

- $\bm\beta = (1,2)$
- $\sigma_{\epsilon}^2 = 0.01$
- $\sigma_{e}^2 = 0.1$
- marginal generalized variance the $s_i$ equal to 1
- $\lambda_1 = 0.5$

We fit a correctly specified model with

- flat priors for each $\beta_c$
- Gamma(1, 0.00005) prior on $1 / \sigma^2_{\epsilon}$
- BYM2 specification of the ICAR and IID random effects
  - PC prior on the total standard deviation such that there is a 1\% chance it is bigger than 2
  - PC prior on the proportion of the variance that is spatial such that there is a 50\% chance it is bigger than 0.5
- N(0, 1) prior on $\lambda$

```{r}
# parameters
nreg <- nrow(admin1.mat)
ncause <- 2
beta <- 1:ncause
sigma_epsilon <- 0.01^0.5
sigma_e <- 0.1^.5
lambda <- 0.5

# simulate data
# nsims <- 10
# for (i in 1:nsims) {
#     
# }
e <- vector(mode = "list", length = ncause)
s <- vector(mode = "list", length = ncause)
epsilon <- vector(mode = "list", length = ncause)
y <- vector(mode = "list", length = ncause)
region <- vector(mode = "list", length = ncause)
for (i in 1:ncause) {
  e[[i]] <- rnorm(nreg, 0, sigma_e)
  s[[i]] <- rst(n = 1, type = "s", type.s = "ICAR", Amat = admin1.mat, scale.model = TRUE)
  epsilon[[i]] <- rnorm(nreg, 0, sigma_epsilon)
  region[[i]] <- c(rep(rep(NA, nreg), i - 1), 1:nreg, rep(rep(NA, nreg), ncause - i))
}

cause <- rep(1:ncause, each = nreg)

if (ncause == 2) {
  regionint <- c(rep(NA, nreg), 1:nreg)
  y[[1]] <- beta[1] + s[[1]] + e[[1]] + epsilon[[1]]
  y[[2]] <- beta[2] + lambda * (s[[1]] + e[[1]]) + s[[2]] + e[[2]] + epsilon[[2]]
} else {
  stop("this number of causes not yet implemented")
}

# priors
fe.prec <- list(prec.intercept = 0,
                prec = 0)
bym2_priors <- list(phi=list(prior="pc", param=c(0.5, 0.5), initial=1), 
                    prec=list(prior="pc.prec", param=c(2, 0.01), initial=5))
hyper_lambda <- list(theta = list(prior = 'normal', param = c(0, 1)))

if (ncause == 2) {
  # model formula
  m.form <- y ~ -1 + factor(cause) + 
      f(region1, model = "bym2", 
        graph = "bgd.graph",
        scale.model = TRUE,
        constr = TRUE,
        hyper = bym2_priors) +
      f(region2, model = "bym2", 
        graph = "bgd.graph",
        scale.model = TRUE,
        constr = TRUE,
        hyper = bym2_priors) +
      f(region12, copy = "region1", 
        fixed = FALSE, hyper = hyper_lambda)
  
  # fit model
  mod1 <- inla(m.form, 
               data = data.frame(y = do.call(c, y), cause = cause, 
                                 region1 = region[[1]],
                                 region2 = region[[2]],
                                 region12 = regionint,
                                 prec_loggaussian = sigma_epsilon),
               family = "gaussian",
               control.fixed = fe.prec,
               control.predictor=list(compute=TRUE),
               control.compute=list(config = TRUE),
               control.family = list(hyper = list(prec = list(initial = log(1), fixed=TRUE))),
               scale = prec_loggaussian)
} else {
  stop("this number of causes not yet implemented")
}

# results
fe.res <- mod1$summary.fixed
re.res <- mod1$summary.random
hyperpar.res <- mod1$summary.hyperpar
fitted.res <- mod1$summary.fitted.values

res.table <- data.frame(parameter = c("$\\beta_1$", "$\\beta_2$", 
                         # "$\\sigma_{\\epsilon}$", 
                         "$\\sigma_{T,1}$", "$\\phi_1$",
                         "$\\sigma_{T,2}$", "$\\phi_2$",
                         "$\\lambda$"),
                        med = c(fe.res$`0.5quant`, 
                                1/(hyperpar.res$`0.5quant`[1])^0.5, 
                                hyperpar.res$`0.5quant`[2],
                                1/(hyperpar.res$`0.5quant`[3])^0.5, 
                                hyperpar.res$`0.5quant`[4:5]),
                        L95 = c(fe.res$`0.025quant`, 
                                1/(hyperpar.res$`0.975quant`[1])^0.5, 
                                hyperpar.res$`0.025quant`[2],
                                1/(hyperpar.res$`0.975quant`[3])^0.5, 
                                hyperpar.res$`0.025quant`[4:5]),
                        U95 = c(fe.res$`0.975quant`, 
                                1/(hyperpar.res$`0.025quant`[1])^0.5, 
                                hyperpar.res$`0.975quant`[2],
                                1/(hyperpar.res$`0.025quant`[3])^0.5, 
                                hyperpar.res$`0.975quant`[4:5]))
kable(res.table)

plot(fitted.res$`0.5quant` ~ do.call(c, y), main = "Comparison of fitted values with 95% UI",
     xlab = "observed", ylab = "fitted",
     ylim = range(c(fitted.res$`0.025quant`, fitted.res$`0.975quant`)))
segments(do.call(c, y), fitted.res$`0.025quant`, do.call(c, y), fitted.res$`0.975quant`)
abline(0,1, col = "red")

# plot(re.res$region1$`0.5quant` ~ gamma1, main = "Comparison of REs with 95% UI",
#      xlab = "observed", ylab = "fitted", 
#      ylim = range(c(re.res$region1$`0.025quant`, re.res$region1$`0.975quant`)))
# segments(gamma1, re.res$region1$`0.025quant`, gamma1, re.res$region1$`0.975quant`)
# abline(0,1, col = "red")
# 
# plot(re.res$region2$`0.5quant` ~ gamma2, main = "Comparison of REs with 95% UI",
#      xlab = "observed", ylab = "fitted", 
#      ylim = range(c(re.res$region2$`0.025quant`, re.res$region2$`0.975quant`)))
# segments(gamma2, re.res$region2$`0.025quant`, gamma2, re.res$region2$`0.975quant`)
# abline(0,1, col = "red")
```

However, this model is sensitive to the variance parameters (for good reason). We next simulate data and fit the same model with the following parameters:

- $\bm\beta = (1,2)$
- $\sigma_{\epsilon}^2 = 0.1$
- $\sigma_{e}^2 = 0.25$
- marginal generalized variance the $s_i$ equal to 1
- $\lambda_1 = 0.5$

```{r}
# parameters
nreg <- nrow(admin1.mat)
ncause <- 2
beta <- 1:ncause
sigma_epsilon <- 0.1^0.5
sigma_e <- 0.25^.5
lambda <- 0.5

# simulate data
# nsims <- 10
# for (i in 1:nsims) {
#     
# }
e <- vector(mode = "list", length = ncause)
s <- vector(mode = "list", length = ncause)
epsilon <- vector(mode = "list", length = ncause)
y <- vector(mode = "list", length = ncause)
region <- vector(mode = "list", length = ncause)
for (i in 1:ncause) {
  e[[i]] <- rnorm(nreg, 0, sigma_e)
  s[[i]] <- rst(n = 1, type = "s", type.s = "ICAR", Amat = admin1.mat, scale.model = TRUE)
  epsilon[[i]] <- rnorm(nreg, 0, sigma_epsilon)
  region[[i]] <- c(rep(rep(NA, nreg), i - 1), 1:nreg, rep(rep(NA, nreg), ncause - i))
}

cause <- rep(1:ncause, each = nreg)

if (ncause == 2) {
  regionint <- c(rep(NA, nreg), 1:nreg)
  y[[1]] <- beta[1] + s[[1]] + e[[1]] + epsilon[[1]]
  y[[2]] <- beta[2] + lambda * (s[[1]] + e[[1]]) + s[[2]] + e[[2]] + epsilon[[2]]
} else {
  stop("this number of causes not yet implemented")
}

# priors
fe.prec <- list(prec.intercept = 0,
                prec = 0)
bym2_priors <- list(phi=list(prior="pc", param=c(0.5, 0.5), initial=1), 
                    prec=list(prior="pc.prec", param=c(2, 0.01), initial=5))
hyper_lambda <- list(theta = list(prior = 'normal', param = c(0, 1)))

if (ncause == 2) {
  # model formula
  m.form <- y ~ -1 + factor(cause) + 
      f(region1, model = "bym2", 
        graph = "bgd.graph",
        scale.model = TRUE,
        constr = TRUE,
        hyper = bym2_priors) +
      f(region2, model = "bym2", 
        graph = "bgd.graph",
        scale.model = TRUE,
        constr = TRUE,
        hyper = bym2_priors) +
      f(region12, copy = "region1", 
        fixed = FALSE, hyper = hyper_lambda)
  
  # fit model
  mod1 <- inla(m.form, 
               data = data.frame(y = do.call(c, y), cause = cause, 
                                 region1 = region[[1]],
                                 region2 = region[[2]],
                                 region12 = regionint),
               family = "gaussian",
               control.fixed = fe.prec,
               control.predictor=list(compute=TRUE),
               control.compute=list(config = TRUE))
} else {
  stop("this number of causes not yet implemented")
}

mod1$summary.fixed
mod1$summary.random$region1
mod1$summary.random$region2
mod1$summary.random$region12
mod1$summary.hyperpar
# mod1$summary.fitted.values$`0.5quant` - y
plot(mod1$summary.fitted.values$`0.5quant` ~ do.call(c, y))
abline(0,1, col = "red")
```

We will isolate the ICAR component and simulate data from the following model:

\eq{
    y_{i1} &= \beta_1 + s_{i1} + \epsilon_{i1} \\
    y_{i2} &= \beta_2 + \lambda_1 (s_{i1}) + s_{i2} + \epsilon_{i2} \\
    \epsilon_{ic} &\overset{IID}{\sim} N(0, \sigma^2_{\epsilon})\\
    \bm{s_c} &\sim ICAR(\sigma_{s}^2) \\
}

First, we simulate data for $i = 1, \dots, 8$ regions and $c = 1, \dots, 2$ causes, with

- $\bm\beta = (1,2)$
- $\sigma_{\epsilon}^2 = 0.1$
- marginal generalized variance the $s_i$ equal to 1
- $\lambda_1 = 0.5$

We fit a correctly specified model with

- flat priors for each $\beta_c$
- Gamma(1, 0.00005) prior on $1 / \sigma^2_{\epsilon}$
- PC prior on the ICAR standard deviation such that there is a 1\% chance it is bigger than 2
- N(0, 1) prior on $\lambda$

```{r}
# parameters
nreg <- nrow(admin1.mat)
ncause <- 2
beta <- 1:ncause
sigma_epsilon <- 0.1^0.5
lambda <- 0.5

# simulate data
# nsims <- 10
# for (i in 1:nsims) {
#     
# }
s <- vector(mode = "list", length = ncause)
epsilon <- vector(mode = "list", length = ncause)
y <- vector(mode = "list", length = ncause)
region <- vector(mode = "list", length = ncause)
for (i in 1:ncause) {
  s[[i]] <- rst(n = 1, type = "s", type.s = "ICAR", Amat = admin1.mat, scale.model = TRUE)
  epsilon[[i]] <- rnorm(nreg, 0, sigma_epsilon)
  region[[i]] <- c(rep(rep(NA, nreg), i - 1), 1:nreg, rep(rep(NA, nreg), ncause - i))
}

cause <- rep(1:ncause, each = nreg)

if (ncause == 2) {
  regionint <- c(rep(NA, nreg), 1:nreg)
  y[[1]] <- beta[1] + s[[1]] + epsilon[[1]]
  y[[2]] <- beta[2] + lambda * (s[[1]]) + s[[2]] + epsilon[[2]]
} else {
  stop("this number of causes not yet implemented")
}

# priors
fe.prec <- list(prec.intercept = 0,
                prec = 0)
besag_prior <- list(prec = list(prior = "pc.prec", param=c(2, 0.01), initial=5))
hyper_lambda <- list(theta = list(prior = 'normal', param = c(0, 1)))

if (ncause == 2) {
  # model formula
  m.form <- y ~ -1 + factor(cause) + 
      f(region1, model = "besag", 
        graph = "bgd.graph",
        scale.model = TRUE,
        constr = TRUE,
        hyper = besag_prior) +
      f(region2, model = "besag", 
        graph = "bgd.graph",
        scale.model = TRUE,
        constr = TRUE,
        hyper = besag_prior) +
      f(region12, copy = "region1", 
        fixed = FALSE, hyper = hyper_lambda)
  
  # fit model
  mod1 <- inla(m.form, 
               data = data.frame(y = do.call(c, y), cause = cause, 
                                 region1 = region[[1]],
                                 region2 = region[[2]],
                                 region12 = regionint),
               family = "gaussian",
               control.fixed = fe.prec,
               control.predictor=list(compute=TRUE),
               control.compute=list(config = TRUE))
} else {
  stop("this number of causes not yet implemented")
}

mod1$summary.fixed
mod1$summary.random$region1
mod1$summary.random$region2
mod1$summary.random$region12
mod1$summary.hyperpar
# mod1$summary.fitted.values$`0.5quant` - y
plot(mod1$summary.fitted.values$`0.5quant` ~ do.call(c, y))
abline(0,1, col = "red")
```

<!-- # For later: multinomial modeling -->

<!-- ## Data generating mechanism -->

<!-- Let $i = 1, \dots, n$ index regions each with population size $N_i$, and $c = 1, \dots, C$ index causes of death. Let $\mathbf{y}_{i} = (y_{i1}, \dots, y_{iC})$ be cause-specific death counts in region $i$. We will generate data such that -->

<!-- $$\mathbf{y}_{i} | \mathbf{\pi}_i \sim \mbox{Multinomial}(N_i, \mathbf{\pi}_i)$$ -->

<!-- ### DGM 1 -->

<!-- $$\mbox{logit}(\pi_{ic}) = \beta_c + s_i + e_i$$ -->

<!-- with $beta_c$ fixed intercepts, $\mathbf{e} \overset{IID}{\sim} N(0, \sigma^2_e)$, and $\mathbf{s} \sim \mbox{ICAR}(\sigma^2_s)$. -->

<!-- #### Simulated data -->

<!-- ```{r} -->
<!-- ## load bgd shape files -->
<!-- poly.adm0 <- readOGR(dsn = "/Users/austin/Dropbox/dissertation_2/survey-csmf/data/shapefiles/bgd_adm_bbs_20201113_SHP", -->
<!--                      layer = "bgd_admbnda_adm0_bbs_20201113") -->
<!-- poly.adm0$region <- "All" -->
<!-- poly.adm1 <- readOGR(dsn = "/Users/austin/Dropbox/dissertation_2/survey-csmf/data/shapefiles/bgd_adm_bbs_20201113_SHP", -->
<!--                      layer = "bgd_admbnda_adm1_bbs_20201113") -->
<!-- poly.adm1$region <- tolower(poly.adm1$ADM1_EN) -->

<!-- # get adjacency matrix -->
<!-- admin1.mat <- poly2nb(SpatialPolygons(poly.adm1@polygons)) -->
<!-- admin1.mat <- nb2mat(admin1.mat, zero.policy = TRUE) -->

<!-- # set parameters -->
<!-- N <- nrow(admin1.mat) -->
<!-- ncause <- 3 -->
<!-- ni <- rep(1000, N) -->
<!-- # simulate data -->
<!-- s <- rst(n = N, type = "s", type.s = "ICAR", Amat = admin1.mat, scale.model = TRUE) -->
<!-- e <- rnorm(N, 0, 0.1) -->
<!-- beta <- rep(1, ncause) -->

<!-- y <- rmultinom(1, ni, beta + e + s) -->

<!-- poly.adm1$y <- y -->
<!-- med.palette <- brewer.pal(n = 5, name = "Purples") -->
<!-- med.int <- classIntervals(round(direct.admin1$logit.est, 2), -->
<!--                           n = 5, style = 'jenks') -->
<!-- med.col <- findColours(med.int, med.palette) -->
<!-- par(mfrow = c(2, 4)) -->
<!-- for(year in periods){ -->
<!--   tmp <- poly.adm1 -->
<!--   for (reg in tmp$region) { -->
<!--       tmp$col[tmp$region == reg] <- med.col[direct.admin1$years == year & direct.admin1$region == reg] -->
<!--   } -->
<!--   plot(tmp, border = F, col = tmp$col, -->
<!--        axes = F, main = year) -->
<!-- } -->
<!-- plot(NA, xlim = c(0,1), ylim = c(0,1), axes = F, xlab = "", ylab = "") -->
<!-- legend(x = "center",inset = 0, -->
<!--        legend = names(attr(med.col, 'table')), -->
<!--        fill = med.palette, cex= 1, horiz = FALSE, bty = 'n') -->

<!-- ``` -->

<!-- ### DGM 2 (coregionalization) -->

<!-- \begin{align*} -->
<!--     \mbox{logit}(\pi_{i1}) &= \beta_1 + s_{i1} + e_i \\ -->
<!--     \mbox{logit}(\pi_{i2}) &= \beta_1 + \lambda_1^2 s_{i1} + s_{i2}+ e_i \\ -->
<!--     \dots & \\ -->
<!--     \mbox{logit}(\pi_{iC}) &= \beta_1 + \lambda_1^C s_{i1} + \lambda_2^C s_{i2} + \dots + s_{iC}+ e_i -->
<!-- \end{align*} -->


