---
title: "Individual-level discrete time competing risks survival modeling"
author: "Austin Schumacher"
date: "`r Sys.Date()`"
output: 
    pdf_document:
        includes: 
            in_header: "/Users/austin/Desktop/work/presentation_publication_resources/mycommands.tex"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE)
```

# Simulation using "fake" data

## Multinomial

### 3 categories

We will first simulate $n$ draws from a multinomial distribution with 3 categories

```{r}
library(devtools)
# install_github("https://github.com/richardli/SUMMER/tree/dev") # Johnny's dev version with pop sim functions
library(svyVGAM)
library(INLA)
library(SUMMER)
library(fields)
library(tidyverse)
library(magrittr)
library(data.table)
library(mvtnorm)

# set seed
set.seed(98125)

# sim setup and storage of results
nsim <- 100
p.res <- matrix(NA, nrow = nsim, ncol = 2)
coverage.res <- matrix(NA, nrow = nsim, ncol = 2)
cov.res <-array(NA, dim = c(2, 2, nsim))

# parameters
n <- 1000
alpha <- c(-2, -1)
denom_p <- sum(exp(alpha)) + 1
p_cause <- c(exp(alpha)/denom_p)
p <- c(p_cause, 1 - sum(p_cause))

for (s in 1:nsim) {
    # cat(paste0("Starting sim ", s, "....\n"))
    # simulate data
    y <- rmultinom(n, 1, p)
    dat <- data.frame(obs = 1:n,
                      cause1 = y[1,],
                      cause_other = y[2,],
                      alive = y[3,])
    
    # fit model
    my.svydesign <- survey::svydesign(ids = ~ 1,
                                      data = dat)
    
    mult.mod <- svy_vglm(cbind(cause1, cause_other, alive) ~ 1, 
                         family = multinomial, 
                         design = my.svydesign)
    denom <- 1 + sum(exp(mult.mod$coef))
    p.res[s, ] <- exp(mult.mod$coef)/denom
    cis <- confint(mult.mod)
    coverage.res[s, 1] <- alpha[1] > cis[1, 1] & alpha[1] < cis[1, 2]
    coverage.res[s, 2] <- alpha[2] > cis[2, 1] & alpha[2] < cis[2, 2]
    cov.res[,,s] <- mult.mod$var
}

boxplot(p.res[, 1], main = "Posterior medians of p1", ylim = range(c(p.res[, 1], p[1])))
abline(h = p[1], col = "red")

boxplot(p.res[, 2], main = "Posterior medians of p2", ylim = range(c(p.res[, 2], p[2])))
abline(h = p[2], col = "red")

apply(coverage.res, 2, mean)

apply(cov.res, c(1, 2), mean)
```

## Individual-level SRS

### 3 categories (1 cause, all other causes, and alive), no age modeling

We will simulate $n$ individuals over 60 months, where in each month for all alive individuals, we apply the multinomial probabilities of dying from cause 1, dying from all other causes, or remaining alive.

Thus, we have the discrete-time survival data generating mechanism as follows:

TODO write this up

```{r}
# sim setup and storage of results
nsim <- 100
p.res <- matrix(NA, nrow = nsim, ncol = 2)
coverage.res <- matrix(NA, nrow = nsim, ncol = 2)
cov.res <-array(NA, dim = c(2, 2, nsim))

# parameters
n <- 1000
alpha <- c(-4, -3)
denom_p <- sum(exp(alpha)) + 1
p_cause <- c(exp(alpha)/denom_p)
p <- c(p_cause, 1 - sum(p_cause))
p_cum <- cumsum(p)

for (s in 1:nsim) {
    # simulate data
    y <- vector(mode = "list", length = n)
    for(i in 1:n) {
        y[[i]] <- matrix(NA, ncol = 5, nrow = 60)
        y[[i]][, 4] <- i # person id
        y[[i]][, 5] <- 1:60 # month indicator
        # cat(paste("\nStarting person", i, "\n"))
        t <- 1
        death <- 0
        while(death == 0) {
            # if (t %% 10 == 0) cat(paste("\t month", t, "\n"))
            rand <- runif(1, 0, 1)
            if (rand < p_cum[1]) {
                y[[i]][t, 1:3] <- c(1, 0, 0)
                death <- 1
            } else if (rand >= p_cum[1] & rand < p_cum[2]) {
                y[[i]][t, 1:3] <- c(0, 1, 0)
                death <- 1
            } else {
                y[[i]][t, 1:3] <- c(0, 0, 1)
            }
            t <- t + 1
            if (t == 60) death <- 1
        }
    }
    
    ymat <- do.call(rbind, y)
    ymat <- ymat[complete.cases(ymat), ]
    
    # create data frame
    dat <- data.frame(id = ymat[, 4],
                      time = ymat[, 5],
                      cause1 = ymat[, 1],
                      cause_other = ymat[, 2],
                      alive = ymat[, 3])
    
    # fit model
    my.svydesign <- survey::svydesign(ids = ~ id,
                                      data = dat)
    
    mult.mod <- svy_vglm(cbind(cause1, cause_other, alive) ~ 1, 
                         family = multinomial, 
                         design = my.svydesign)
    denom <- 1 + sum(exp(mult.mod$coef))
    p.res[s, ] <- exp(mult.mod$coef)/denom
    cis <- confint(mult.mod)
    coverage.res[s, 1] <- alpha[1] > cis[1, 1] & alpha[1] < cis[1, 2]
    coverage.res[s, 2] <- alpha[2] > cis[2, 1] & alpha[2] < cis[2, 2]
    cov.res[,,s] <- mult.mod$var
}

# results
boxplot(p.res[, 1], main = "Posterior medians of p1", ylim = range(c(p.res[, 1], p[1])))
abline(h = p[1], col = "red")

boxplot(p.res[, 2], main = "Posterior medians of p2", ylim = range(c(p.res[, 2], p[2])))
abline(h = p[2], col = "red")

apply(coverage.res, 2, mean)

apply(cov.res, c(1, 2), mean)
```

### 3 categories (1 cause, all other causes, and alive), age-group-specific intercepts

We will again simulate $n$ individuals over 60 months, where in each month for all alive individuals, we apply the multinomial probabilities of dying from cause 1, dying from all other causes, or remaining alive. This time, those probabilities are allowed to vary in 6 different age bands.

```{r}
# sim setup and storage of results
nsim <- 50
alpha.res <- matrix(NA, nrow = nsim, ncol = 12)
coverage.res <- matrix(NA, nrow = nsim, ncol = 12)
cov.res <-array(NA, dim = c(12, 12, nsim))

# parameters
n <- 1000
alpha <- matrix(c(seq(-4, -6, length.out = 6), seq(-3, -5.5, length.out = 6)), 
                nrow = 6, ncol = 2)
alpha_vec <- as.vector(alpha)
denom_p <- apply(alpha, 2, function(x){sum(exp(x)) + 1})
p_cause <- cbind(exp(alpha[,1])/denom_p[1], exp(alpha[,2])/denom_p[2])
p <- cbind(p_cause, 1 - apply(p_cause, 1, sum))
p_cum <- apply(p, 1, cumsum)
p_cum_monthly <- p_cum[,c(1, rep(2,11), rep(c(3,4,5,6), each = 12))]
    
# simulate data
for (s in 1:nsim) {
    # simulate data
    y <- vector(mode = "list", length = n)
    for(i in 1:n) {
        y[[i]] <- matrix(NA, ncol = 6, nrow = 60)
        y[[i]][, 4] <- i # person id
        y[[i]][, 5] <- 1:60 # month indicator
        y[[i]][, 6] <- c(1, rep(2,11), rep(c(3,4,5,6), each = 12)) # age group indicator
        # cat(paste("\nStarting person", i, "\n"))
        t <- 1
        death <- 0
        while(death == 0) {
            # if (t %% 10 == 0) cat(paste("\t month", t, "\n"))
            rand <- runif(1, 0, 1)
            if (rand < p_cum_monthly[1, t]) {
                y[[i]][t, 1:3] <- c(1, 0, 0)
                death <- 1
            } else if (rand >= p_cum_monthly[1, t] & rand < p_cum_monthly[2, t]) {
                y[[i]][t, 1:3] <- c(0, 1, 0)
                death <- 1
            } else {
                y[[i]][t, 1:3] <- c(0, 0, 1)
            }
            t <- t + 1
            if (t == 60) death <- 1
        }
    }
    
    ymat <- do.call(rbind, y)
    ymat <- ymat[complete.cases(ymat), ]
    
    # create data frame
    dat <- data.frame(id = ymat[, 4],
                      time = ymat[, 5],
                      cause1 = ymat[, 1],
                      cause_other = ymat[, 2],
                      alive = ymat[, 3],
                      agegp = ymat[, 6])
    
    # fit model
    my.svydesign <- survey::svydesign(ids = ~ id,
                                      data = dat)
    
    mult.mod <- svy_vglm(cbind(cause1, cause_other, alive) ~ -1 + factor(agegp), 
                         family = multinomial, 
                         design = my.svydesign)
    alpha.res[s,] <- coef(mult.mod)[c(1,3,5,7,9,11,2,4,6,8,10,12)]
    cis <- confint(mult.mod)[c(1,3,5,7,9,11,2,4,6,8,10,12),]
    coverage.res[s, ] <- alpha_vec > cis[,1] & alpha_vec < cis[,2]
    cov.res[,,s] <- mult.mod$var
}

# results
boxplot(alpha.res, use.cols= TRUE, main = "Posterior medians of alpha parameters")
points(1:12, alpha_vec, col = "red", pch = 19)

apply(coverage.res, 2, mean)

apply(cov.res, c(1, 2), mean)
```

# Simulation with Kenya-like data

First, we will simulate a population of births in enumeration areas (point locations) using an inhomogeneous Poisson process with rate proportional to population density in Kenya (using data from IPUMS). 

TODO extract this same info for BGD.

EAs and populations are dispersed conditional on the "known number" of EAs, households, and target population at the admin 1 level using multilevel multinomial sampling, first sampling the EAs, then distributing households among the EAs, then the target population among the households. Any areal level below the 'areas' we call 'subareas'. For instance, the 'areas' might be Admin1 if that is the smallest level at which the number of EAs, households, and people is known, and the 'subareas' might be Admin2. The multilevel multinomial sampling may be stratified by urban/rural within the areas if the number of EAs, households, and people is also approximately known at that level.

## Each region has it's own probability of selection

### 2 outcomes (dead or alive): Binomial model

WRITE OUT THE MODEL

## Each region has it's own probability of selection

Simulate a population

```{r}
# simulate population similar to kenya
data(kenyaMaps)
data(kenyaPopulationData)
pixelCoords <- cbind(popMatKenya$east, popMatKenya$north)
SPDEArgs <- list(coords=pixelCoords, nsim=1, margVar=1, effRange=400, 
                mesh=kenyaMesh, inla.seed=12L)
uDraws <- do.call("simSPDE", SPDEArgs)

# simulate the enumeration areas
sigmaEpsilonDraws <- 0.1

# simulate EA and pixel level populations given the risk 
# simulated by the SPDE model
pixelPop <- simPopPixel(uDraws=uDraws, sigmaEpsilonDraws=sigmaEpsilonDraws, 
                       easpa=easpaKenyaNeonatal, 
                       popMat=popMatKenya, targetPopMat=popMatKenyaNeonatal, 
                       stratifyByUrban=TRUE, 
                       doFineScaleRisk=TRUE, poppsub=poppsubKenya, 
                       subareaLevel=TRUE, min1PerSubarea=TRUE, 
                       returnEAinfo=TRUE)

# get data with row for each birth
eadatlong <- as.data.frame(lapply(pixelPop$eaDatList[[1]], rep, eadat$N))
eadatlong$reg <- as.numeric(eadatlong$area)
eadatlong <- eadatlong[order(eadatlong$reg),]

# apply model to get deaths
set.seed(128)
nageg <- 6
nreg <- length(unique(eadatlong$reg))
n <- nrow(eadatlong)
agegroup.probs <- c(0.04, 0.004, rep(0.001, nageg - 2))
alpha_vec <- logitlink(agegroup.probs)

# gamma <- rnorm(nreg, 0, 0.5)
gamma <- rep(0, nreg)

# get probs of death
probs <- data.table(reg = rep(1:nreg, each = nageg),
                    ageg = rep(1:nageg, nreg),
                    alpha = rep(alpha_vec, nreg),
                    gamma = rep(gamma, each = nageg))
probs %<>% mutate(eta = alpha + gamma) %>%
    mutate(prob = expit(eta))

# multinomial draws
ns <- c(1, 11, 12, 12, 12, 12) # number of months for each age group
monthly.agegroups <- rep(1:6, times = ns)
agegroup.ref <- c(monthly.agegroups, nageg + 1)
outcome.month <- vector(mode = "list", length = nreg)
outcome.death <- vector(mode = "list", length = nreg)
for (i in 1:nreg) {
    cat(paste0("region: ", i, "\n"))
    tmp <- eadatlong %>% filter(reg == i)
    tmp_probs <- probs %>% 
        filter(reg == i)
    tmp_probs$nmonths <- rep(ns, ncause)
    mult_probs <- rep(tmp_probs$prob, times = tmp_probs$nmonths)
    mult_probs <- c(mult_probs, 1 - sum(mult_probs))
    y.draws <- rmultinom(nrow(tmp), 1, mult_probs)
    y.multinom.category <- apply(y.draws, 2, function(x) {which(x == 1)})
    outcome.month[[i]] <- ifelse(y.multinom.category %in% 1:sum(ns), 
                            y.multinom.category, 
                            y.multinom.category - sum(ns))
    outcome.cause[[i]] <- multinom.cause.ref[y.multinom.category]
}
eadatlong$out.month <- do.call(c, outcome.month)
eadatlong$out.cause <- do.call(c, outcome.cause)

# true 5q0^c
outcome.5q0.c <- prop.table(table(eadatlong$out.cause))[1:ncause]

# different prob weights for each region
avg_prob_samp <- 0.001
reg.p.weight <- expit(rnorm(nreg, logitlink(avg_prob_samp), 0.1))
eadatlong$pweight <- reg.p.weight[eadatlong$reg]
```

Simulate a DHS, format data, run model

```{r}
# function to calculate 5q0^c from multinomial model
get_5q0 <- function(beta, n) {
    ## For testing
    # beta <- coef(mult.mod)
    # n <- c(1, 11, 12, 12, 12, 12)
    
    betas_of_interest <- beta[seq(1,length(beta), by = 2)]
    betas_other <- beta[seq(2,length(beta), by = 2)]
    
    gamma_j <- (1 + exp(betas_of_interest) + exp(betas_other))^-1
    sum_nj_of_gammajs <- rep(0, length(n))
    for (j in 1:length(n)) {
        for (s in 1:n[j]) {
            sum_nj_of_gammajs[j] <- sum_nj_of_gammajs[j] + gamma_j[j]^s
        }
    }
    
    phi_of_interest <- sum(exp(betas_of_interest) * sum_nj_of_gammajs)
    phi_other <- sum(exp(betas_other) * sum_nj_of_gammajs)
    
    return(c(phi_of_interest, phi_other))
}

# sim setup and storage of results
set.seed(2008)
nsim <- 100
q5.res <- matrix(NA, nrow = nsim, ncol = ncause)
coverage.res <- matrix(NA, nrow = nsim, ncol = ncause)

# simulate data
for (s in 1:nsim) {
    
    # testing
    # s <- 1
    
    cat(paste0("Starting sim ", s, "... "))
    
    # Sample births
    ps <- runif(nrow(eadatlong), 0, 1)
    sampdat <- eadatlong[ps < eadatlong$pweight,]

    # simulate data
    cat(paste0("formatting data... "))
    
    # expand to monthly data
    sampdat$id <- 1:nrow(sampdat)
    sampdat$out.month[sampdat$out.month == 61] <- 60
    finalmonths <- sampdat$out.month
    sampdat.long <- sampdat[rep(1:nrow(sampdat), times = sampdat$out.month),]
    month.long <- unlist(lapply(sampdat$out.month, function(x) {1:x}))
    sampdat.long$month <- month.long
    agegroup.month.ref <- rep(1:nageg, times = ns)
    sampdat.long$ageg <- agegroup.month.ref[sampdat.long$month]
    sampdat.long$cause1 <- ifelse(sampdat.long$month == sampdat.long$out.month & 
                                      sampdat.long$out.cause == 1,
                                  1, 0)
    sampdat.long$cause_other <- ifelse(sampdat.long$month == sampdat.long$out.month & 
                                           sampdat.long$out.cause == 2,
                                       1, 0)
    sampdat.long$alive <- ifelse(sampdat.long$cause1 == 0 & 
                                       sampdat.long$cause_other == 0,
                                 1, 0)
    
    # compactify
    sampdat.comp <- sampdat.long[, c("reg", "pweight", "ageg",
                                     "alive", "cause1", "cause_other")]
    sampdat.comp$total <- 1
    formula <- as.formula(".~ reg + ageg + pweight")
    sampdat.comp <- aggregate(formula, data = sampdat.comp, FUN = sum, drop = TRUE)
    # table(sampdat.comp$age)
    
    # fit model
    cat(paste0("fitting model stage 1 (survey multinomial) model... "))
    
    # survey design
    my.svydesign <- survey::svydesign(ids = ~ reg, weights = ~ pweight, 
                                      data = sampdat.comp)
    
    ## multinomial model
    mult.mod <- svy_vglm(cbind(cause1, cause_other, alive) ~ -1 + factor(ageg), 
                         family = multinomial, 
                         design = my.svydesign)
    
    # extract betas and sigma-hat to use in calculating 5q0 and its variance
    betas <- coef(mult.mod)
    V <- stats::vcov(mult.mod)
    
    #  simulate betas
    betasim <- rmvnorm(10000, mean = betas, sigma = V)
    q5.sim <- t(apply(betasim, 1, function(x) get_5q0(x, ns)))
    bounds <- apply(q5.sim, 2, quantile, c(0.025, 0.975))
    
    q5.res[s, ] <- get_5q0(betas, ns)
    coverage.res[s, ] <- c(bounds[1, 1] < outcome.5q0.c[1] & bounds[2, 1] > outcome.5q0.c[1],
                           bounds[1, 2] < outcome.5q0.c[2] & bounds[2, 2] > outcome.5q0.c[2])
    
    cat(paste0("done.\n"))
}

# results
apply(coverage.res, 2, mean)
```


### 3 outcomes (cause 1, other causes, alive): Multinomial model

Simulate a population

```{r}
# simulate population similar to kenya
data(kenyaMaps)
data(kenyaPopulationData)
pixelCoords <- cbind(popMatKenya$east, popMatKenya$north)
SPDEArgs <- list(coords=pixelCoords, nsim=1, margVar=1, effRange=400, 
                mesh=kenyaMesh, inla.seed=12L)
uDraws <- do.call("simSPDE", SPDEArgs)

# simulate the enumeration areas
sigmaEpsilonDraws <- 0.1

# simulate EA and pixel level populations given the risk 
# simulated by the SPDE model
pixelPop <- simPopPixel(uDraws=uDraws, sigmaEpsilonDraws=sigmaEpsilonDraws, 
                       easpa=easpaKenyaNeonatal, 
                       popMat=popMatKenya, targetPopMat=popMatKenyaNeonatal, 
                       stratifyByUrban=TRUE, 
                       doFineScaleRisk=TRUE, poppsub=poppsubKenya, 
                       subareaLevel=TRUE, min1PerSubarea=TRUE, 
                       returnEAinfo=TRUE)

# get data with row for each birth
eadatlong <- as.data.frame(lapply(pixelPop$eaDatList[[1]], rep, eadat$N))
eadatlong$reg <- as.numeric(eadatlong$area)
eadatlong <- eadatlong[order(eadatlong$reg),]

# apply model to get deaths
set.seed(128)
nageg <- 6
ncause <- 2
nreg <- length(unique(eadatlong$reg))
n <- nrow(eadatlong)
c.frac <- 0.3 # cause fraction for the cause of interest
agegroup.probs <- c(0.04, 0.004, rep(0.001, nageg - 2))
alpha <- matrix(c(logitlink(agegroup.probs * c.frac), logitlink(agegroup.probs * (1 - c.frac))),
                nrow = nageg, ncol = ncause)
alpha_vec <- as.vector(alpha)

# gamma <- rnorm(nreg, 0, 0.5)
gamma <- rep(0, nreg)

# get probs of death
probs <- data.table(reg = rep(1:nreg, each = nageg*ncause),
                    ageg = rep(1:nageg, nreg*ncause),
                    cause = rep(rep(1:ncause, each = nageg), nreg),
                    alpha = rep(alpha_vec, nreg),
                    gamma = rep(gamma, each = nageg*ncause))
probs %<>% mutate(eta = alpha + gamma) %>%
    group_by(cause, reg) %>%
    mutate(prob = exp(eta)/(1 + sum(exp(eta))))

# multinomial draws
ns <- c(1, 11, 12, 12, 12, 12) # number of months for each age group
monthly.agegroups <- rep(1:6, times = ns)
multinom.agegroup.ref <- c(rep(monthly.agegroups, ncause), nageg + 1)
multinom.cause.ref <- c(rep(1:ncause, each = sum(ns)), ncause + 1)
outcome.month <- vector(mode = "list", length = nreg)
outcome.cause <- vector(mode = "list", length = nreg)
for (i in 1:nreg) {
    cat(paste0("region: ", i, "\n"))
    tmp <- eadatlong %>% filter(reg == i)
    tmp_probs <- probs %>% 
        filter(reg == i)
    tmp_probs$nmonths <- rep(ns, ncause)
    mult_probs <- rep(tmp_probs$prob, times = tmp_probs$nmonths)
    mult_probs <- c(mult_probs, 1 - sum(mult_probs))
    y.draws <- rmultinom(nrow(tmp), 1, mult_probs)
    y.multinom.category <- apply(y.draws, 2, function(x) {which(x == 1)})
    outcome.month[[i]] <- ifelse(y.multinom.category %in% 1:sum(ns), 
                            y.multinom.category, 
                            y.multinom.category - sum(ns))
    outcome.cause[[i]] <- multinom.cause.ref[y.multinom.category]
}
eadatlong$out.month <- do.call(c, outcome.month)
eadatlong$out.cause <- do.call(c, outcome.cause)

# true 5q0^c
outcome.5q0.c <- prop.table(table(eadatlong$out.cause))[1:ncause]

# different prob weights for each region
avg_prob_samp <- 0.001
reg.p.weight <- expit(rnorm(nreg, logitlink(avg_prob_samp), 0.1))
eadatlong$pweight <- reg.p.weight[eadatlong$reg]
```

Simulate a DHS, format data, run model

```{r}
# function to calculate 5q0^c from multinomial model
get_5q0 <- function(beta, n) {
    ## For testing
    # beta <- coef(mult.mod)
    # n <- c(1, 11, 12, 12, 12, 12)
    
    betas_of_interest <- beta[seq(1,length(beta), by = 2)]
    betas_other <- beta[seq(2,length(beta), by = 2)]
    
    gamma_j <- (1 + exp(betas_of_interest) + exp(betas_other))^-1
    sum_nj_of_gammajs <- rep(0, length(n))
    for (j in 1:length(n)) {
        for (s in 1:n[j]) {
            sum_nj_of_gammajs[j] <- sum_nj_of_gammajs[j] + gamma_j[j]^s
        }
    }
    
    phi_of_interest <- sum(exp(betas_of_interest) * sum_nj_of_gammajs)
    phi_other <- sum(exp(betas_other) * sum_nj_of_gammajs)
    
    return(c(phi_of_interest, phi_other))
}

# sim setup and storage of results
set.seed(2008)
nsim <- 100
q5.res <- matrix(NA, nrow = nsim, ncol = ncause)
coverage.res <- matrix(NA, nrow = nsim, ncol = ncause)

# simulate data
for (s in 1:nsim) {
    
    # testing
    # s <- 1
    
    cat(paste0("Starting sim ", s, "... "))
    
    # Sample births
    ps <- runif(nrow(eadatlong), 0, 1)
    sampdat <- eadatlong[ps < eadatlong$pweight,]

    # simulate data
    cat(paste0("formatting data... "))
    
    # expand to monthly data
    sampdat$id <- 1:nrow(sampdat)
    sampdat$out.month[sampdat$out.month == 61] <- 60
    finalmonths <- sampdat$out.month
    sampdat.long <- sampdat[rep(1:nrow(sampdat), times = sampdat$out.month),]
    month.long <- unlist(lapply(sampdat$out.month, function(x) {1:x}))
    sampdat.long$month <- month.long
    agegroup.month.ref <- rep(1:nageg, times = ns)
    sampdat.long$ageg <- agegroup.month.ref[sampdat.long$month]
    sampdat.long$cause1 <- ifelse(sampdat.long$month == sampdat.long$out.month & 
                                      sampdat.long$out.cause == 1,
                                  1, 0)
    sampdat.long$cause_other <- ifelse(sampdat.long$month == sampdat.long$out.month & 
                                           sampdat.long$out.cause == 2,
                                       1, 0)
    sampdat.long$alive <- ifelse(sampdat.long$cause1 == 0 & 
                                       sampdat.long$cause_other == 0,
                                 1, 0)
    
    # compactify
    sampdat.comp <- sampdat.long[, c("reg", "pweight", "ageg",
                                     "alive", "cause1", "cause_other")]
    sampdat.comp$total <- 1
    formula <- as.formula(".~ reg + ageg + pweight")
    sampdat.comp <- aggregate(formula, data = sampdat.comp, FUN = sum, drop = TRUE)
    # table(sampdat.comp$age)
    
    # fit model
    cat(paste0("fitting model stage 1 (survey multinomial) model... "))
    
    # survey design
    my.svydesign <- survey::svydesign(ids = ~ reg, weights = ~ pweight, 
                                      data = sampdat.comp)
    
    ## multinomial model
    mult.mod <- svy_vglm(cbind(cause1, cause_other, alive) ~ -1 + factor(ageg), 
                         family = multinomial, 
                         design = my.svydesign)
    
    # extract betas and sigma-hat to use in calculating 5q0 and its variance
    betas <- coef(mult.mod)
    V <- stats::vcov(mult.mod)
    
    #  simulate betas
    betasim <- rmvnorm(10000, mean = betas, sigma = V)
    q5.sim <- t(apply(betasim, 1, function(x) get_5q0(x, ns)))
    bounds <- apply(q5.sim, 2, quantile, c(0.025, 0.975))
    
    q5.res[s, ] <- get_5q0(betas, ns)
    coverage.res[s, ] <- c(bounds[1, 1] < outcome.5q0.c[1] & bounds[2, 1] > outcome.5q0.c[1],
                           bounds[1, 2] < outcome.5q0.c[2] & bounds[2, 2] > outcome.5q0.c[2])
    
    cat(paste0("done.\n"))
}

# results
apply(coverage.res, 2, mean)
```

## DHS-like survey: stratified (by urban-rural) cluster sampling
