---
title: 'Stage 1 modeling simulation to assess coverage: testing the binomial case'
author: "Austin Schumacher"
date: "`r Sys.Date()`"
output: 
    pdf_document:
        toc: true
        toc_depth: 4
        includes: 
            in_header: "/Users/austin/Desktop/work/presentation_publication_resources/mycommands.tex"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache = TRUE)
library(tidyverse); library(survey); library(svyVGAM); library(VGAM); library(magrittr);
library(Hmisc); library(SUMMER); library(mvtnorm);
```

# Binomial modeling

## Data generating mechanism

For 100,000 women, the number of births were generated from a Poisson distribution with rate of 3. We then generate deaths for these children using the following process.

Let $i = 1, \dots, N = 100000$ index individuals, let $t = 0, \dots, T = 59$ index time in months, and define $T_i$ the time of observation for individual $i$. Divide the $60$ months into $J = 6$ age groups indexed by $j[t]$ such that 

$$j[t] = \begin{cases}
1 \text{ if } t = 0 \\
2 \text{ if } 1 \leq t \leq 11 \\
3 \text{ if } 12 \leq t \leq 23 \\
4 \text{ if } 24 \leq t \leq 35 \\
5 \text{ if } 36 \leq t \leq 47 \\
6 \text{ if } 48 \leq t \leq 59
\end{cases}$$

For each individual, we will run Bernoulli trials starting at time $t = 0$ until the first death. These trials will have different probabilities of death for each age group, $\pi_{j[t]}$. The time of death for individual $i$ will be $T_i$. All Bernoulli trials are independent. We will specify regression parameters for a logistic regression model to generate the data. Thus, for $i = 1, \dots, N$ and $t = 1, \dots, T_i$, we have

\eq{
    y_{it} | \pi_{it} &\sim \mbox{Bernoulli}(\pi_{it}) \\
    \mbox{logit}(\pi_{it}) &= \beta_{j[t]}
}

For all models, we will organize the data such that each observation is a person-month, so individual $i$ will contribute a sequence of $T_i - 1$ 0's followed by a single 1 for month $T_i$ (unless no death was observed, in which case there will be all 0's), and each of these 0s and 1s will be its own row in the data set.

We will set the $\beta_{j[t]}$ parameters such that the age-group-specific monthly probabilities $(\mbox{expit}(\beta_1), \dots, \mbox{expit}(\beta_6)) = (0.04, 0.005, 0.004, 0.003, 0.002, 0.001)$.

Here is the code to generate the population:

```{r generation-bin}
# function to calculate 5q0^c from multinomial model
get_5q0_bin <- function(beta, n) {
    ## For testing
    # beta <- coef(bin.mod)
    # n <- c(1, 11, 12, 12, 12, 12)
    
    probs <- expit(beta)
    mean.est <- (1 - prod((1 - probs)^n, na.rm = TRUE))
    return(mean.est)
}

# params
Nw <- 100000 # number of women
Nc <- 500 # number of clusters
nc <- 25 # number of sampled clusters
nw <- 30 # number of sampled women in each cluster
birth_lambda <- 3
nbirths <- rpois(Nw, birth_lambda)
clusterid <- sort(sample(1:Nc, Nw, replace = TRUE))
Nwc <- as.vector(table(clusterid))
weights <- (Nc/nc) * (Nwc/nw)
clusterprobs <- 1/weights

# create data set for population where each row is a birth 
# and simulate deaths
dat <- tibble(motherid = 1:Nw,
              clusterid = clusterid) %>%
    group_by(clusterid) %>%
    mutate(withinclustermotherid = sequence(n()))
dat <- dat[rep(1:Nw, times = nbirths),]
weights.dat <- tibble(clusterid = 1:Nc,
                      weight = weights,
                      prob = clusterprobs)
dat %<>% left_join(weights.dat, by = "clusterid") %>%
    group_by(motherid) %>%
    mutate(birthid = sequence(n())) %>%
    ungroup()  %>%
    mutate(id = 1:n(),
           clustermotherbirthid = paste(clusterid, motherid, birthid, sep = "_"),
           clustermotherid = paste(clusterid, withinclustermotherid, sep = "_"))

# simulate deaths
N <- nrow(dat)
T <- 60
J <- 6

probs <- c(0.04, 0.005, 0.004, 0.003, 0.002, 0.001)
betas <- logitlink(probs)
ns <- c(1, 11, 12, 12, 12, 12)
pis <- rep(probs, times = ns)

dat <- dat[rep(1:N, each = T + 1), ]
dat$t <- rep((0:T), N)
dat$a <- rep(rep(1:(J + 1), times = c(ns, 1)), N)
dat$y <- rbinom((T+1) * N, 1, rep(c(pis, 1), N))
dat %<>% group_by(id) %>%
    filter(cumsum(cumsum(y == 1)) <= 1L) %>%
    ungroup() %>%
    filter(a != 7)

# get 5q0
dat_ind <- dat %>% group_by(id) %>%
    slice_tail(n = 1)
true5q0 <- mean(dat_ind$y)
```

## Cluster sampling coverage simulation

After we have this population of children, we will perform $S = 100$ simulations where we perform a two stage cluster sampling design, fit a logistic regression model to estimate monthly probabilities of death, calculate a 95\% CI for estimated $\nqx{5}{0}$ based on our design-based asymptotic variance estimator, and assess whether it covers the true value of $\nqx{5}{0}$ from the population. 

At the first stage of our two-stage cluster sampling design, $n_c$ clusters were randomly selected from the $N_c$ available. At the second stage, suppose cluster $c$ is selected, then $n_W$ women were randomly selected from the $N_{Wc}$ total women within the selected cluster. The resulting sampling weight for a woman in cluster $c$ is

$$w_{Ec} = \frac{N_c}{n_c} \times \frac{N_{Wc}}{n_W}$$

We set $N_c = 500$, $n_c = 25$, and $n_W = 30$. Note that $N_{Wc}$ is random due to the random allocation of the 100,000 women to the 500 clusters. 

We will then fit a logistic regression model with $J$ age-group-specific intercepts and no overall intercept, accounting for the survey design using the `survey` package in R. We will then extract the covariance matrix and use the delta method to calculate an asymptotic design-based variance estimate. We will use this to calculate a 95\% CI and assess whether it covers the true population $\nqx{5}{0}$.

Here is the code for simulation:

```{r simulation-bin}
# set seed
set.seed(69)

# parameters
nsim <- 100
ns <- c(1, 11, 12, 12, 12, 12)

# results storage
coverage.delta <- rep(NA, nsim)
coverage.jack <- rep(NA, nsim)

# start simulations
for (s in 1:nsim) {
    # testing
    # s <- 1
    
    # sample selection
    sampled_clusters <- sample(1:Nc, nc, replace = FALSE)
    sampled_women <- vector(mode = "list", length = nc)
    for (c in 1:length(sampled_clusters)) {
        sampled_women[[c]] <- sample(1:Nwc[sampled_clusters[c]], nw, replace = FALSE)
    }
    sampled_ids <- tibble(clusterid = rep(sampled_clusters, each = nw),
                          withinclustermotherid = unlist(sampled_women)) %>%
        mutate(clustermotherid = paste(clusterid, withinclustermotherid, sep = "_"))
    dat.sampled <- dat %>% filter(clustermotherid %in% sampled_ids$clustermotherid)

    # survey design
    my.svydesign <- survey::svydesign(ids = ~ clusterid, 
                                      strata = NULL, weights = ~ weight, 
                                      data = dat.sampled)
    
    ## binomial model
    bin.mod <- svyglm(y ~ -1 + factor(a), rescale = TRUE,
                      family = quasibinomial(),
                      design = my.svydesign)
    betas <- coef(bin.mod)
    V <- vcov(bin.mod)
    
    # estimates with delta method var
    mean.est <-get_5q0_bin(betas, ns)
    logit.mean.est <- logitlink(mean.est)
    
    ## partial derivatives ##
    gamma <- prod((1 + exp(betas))^ns, na.rm = TRUE)
    derivatives <- (gamma)/(gamma - 1) * ns * expit(betas)
    
    ## asymptotic var
    var.est <- t(derivatives) %*% V %*% derivatives
    bounds.delta <- expit(rep(logit.mean.est, 2) + rep(var.est^0.5, 2)*qnorm(c(0.025, 0.975)))
    coverage.delta[s] <- (true5q0 > bounds.delta[1]) & (true5q0 < bounds.delta[2])
    
    ## jackknife estimator
    jack5q0s <- rep(NA, nc)
    clusts <- unique(dat.sampled$clusterid)
    for (cc in 1:nc) {
        clust <- clusts[cc]
        tmp <- dat.sampled[dat.sampled$clusterid != clust,]
        
        my.svydesign.tmp <- survey::svydesign(ids = ~ clusterid, 
                                              strata = NULL, weights = ~ weight, 
                                              data = tmp)
    
        ## binomial model
        bin.mod.tmp <- svyglm(y ~ -1 + factor(a), rescale = TRUE,
                              family = quasibinomial(),
                              design = my.svydesign.tmp)
        betas <- coef(bin.mod.tmp)
        
        # estimates
        jack5q0s[cc] <-get_5q0_bin(betas, ns)
    }
    Vjack <- ((nc - 1)/nc) * sum((jack5q0s - mean.est)^2)
    bounds.jack <- rep(mean.est, 2) + rep(Vjack^0.5, 2)*qnorm(c(0.025, 0.975))
    coverage.jack[s] <- (true5q0 > bounds.jack[1]) & (true5q0 < bounds.jack[2])
}
mean(coverage.delta)
mean(coverage.jack)
```
