---
title: "Stage 2 simulation results"
author: "Austin Schumacher"
date: "`r Sys.Date()`"
output: 
    pdf_document:
        extra_dependencies: ["bm", "bbm", "tensor", "pdfpages"]
        includes:
            in_header: "~/Desktop/work/presentation_publication_resources/mycommands.tex"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(SUMMER)
library(tidyverse)
library(spdep)
library(geosphere)
library(haven)
library(knitr)
library(kableExtra)
library(magrittr)
library(rgdal)
library(viridis)
library(classInt)
library(gridExtra)
library(ggpubr)

# load results run info
res_run <- read_csv("results-run-info.csv")
dgm_info <- read_csv("dgm-info.csv")
model_names <- c("Direct estimates", 
                 "Univariate IID",
                 "Univariate BYM",
                 "Bivariate nonshared IID", 
                 "Bivariate nonshared BYM",
                 "Bivariate shared IID", 
                 "Bivariate shared BYM")

dropbox_dir <- "~/Dropbox/dissertation_2/survey-csmf/results/stage-2-simulation"
```

# Overview

We will perform a simulation study to explore the recovery of parameters

# Data generating mechanism

Let $r = 1, \dots, R$ index region and $c = 1, 2$ index outcomes. The region-level means will be generated as

\eq{
    \mu_{r1} &= \beta_1 + v_{r1} + u_{r1} + \lambda (v_{r2} + u_{r2}) \\
    \mu_{r2} &= \beta_2 + v_{r2} + u_{r2} \\
    v_{r1} | \sigma^2_{1} &\overset{iid}{\sim} N(0, \sigma^2_{1}) \\
    v_{r2} | \sigma^2_{2} &\overset{iid}{\sim} N(0, \sigma^2_{2}) \\
    \bm{u}_1 &\sim ICAR(1) \\
    \bm{u}_2 &\sim ICAR(1) 
}
with $\beta_c$ set to be the estimated intercept and $\sigma^2_c$ set to be the estimated variances of the random effects from a model fit to the 2014 KDHS HAZ and WAZ data. The random effects will only be simulated once and the same ones will be used across all simulations.

We will use a BYM2 parameterization for the spatial and IID random effects in each model such that, for $c = 1, 2$, we let

\eq{
    v_{rc} + u_{rc} = \sigma_c(\sqrt{1 - \rho_c}v_{cr}^* + \sqrt{\rho_c}u_{cr}^*)
}

where $v_{cr}^* \overset{iid}{\sim} N(0,1)$ is the unstructured random effect with fixed standard deviation 1 and $\bm{u}_{r}^*$ is the ICAR model scaled so $\mbox{Var}(u_{cr}^*) \approx 1$, which is done by scaling the model so the geometric mean of these variances is 1, using the adjacency matrix to calculate the inverse precision of the ICAR model as in Riebler et al. (2016).

Once we have these $\mu_{rc}$ values, we will simulate the area-level sample means as

\eq{
    \hat{\bm{y}}_{r} | \bm\mu_{r}, \bm{V}_r &\sim N_2(\bm\mu_r, \bm{V}_r)
}

where the $\bm{V}_r$ are set to be equal to the estimated asymptotic design-based covariance matrix of the area-level mean HAZ and WAZ from the 2014 KDHS data.

Now, for an SRS, the sampling distribution of the area-level sample covariance matrices is

\eq{
    (n_r - 1) \hat{\bm{V}}^{srs}_r | V_r, n_r \sim \text{Wishart}\left( \bm{V}_r, (n_r-1) \right) 
}

where $n_r$ is the sample size in area $r$. The 2014 KDHS uses a stratified cluster design rather than an SRS, which has a different sampling variance than an SRS. The ratio of the variance for a statistic calculated using a specific survey design to the variance of that statistic calculated using an SRS of the same sample size is called the design effect,

\eq{
    d^2 &= \frac{\hat{V}}{\hat{V}^{srs}}.
}

In the a typical DHS, the average design effect across all indicators is $d^2 = 1.5^2 = 2.25$ (taken from this answer on the DHS user forum: https://userforum.dhsprogram.com/index.php?t=msg&goto=3448&S=Google). Then, the effective sample size in the stratified design is approximately equal to $n^*_r = n_r/(d^2)$, since the sample variance is linear with respect to the sample size. Thus, we calculate $n^*_r$ as the observed sample size in region $r$ divided by 2.25 and use this to adjust for the survey design effect, giving us

\eq{
    \hat{\bm{V}}_r | \bm{V}_r, n^*_r \sim \frac{1}{(n^*_r - 1)}\text{Wishart}\left( \bm{V}_r, (n^*_r - 1) \right).
}

In each simulation, we will have the same values of $\mu_r$ and $V_r$, and we use these to simulate $\hat{\bm{y}}_r$ and $\hat{\bm{V}}_r$. We will treat the $\hat{\bm{y}}_r$ as pseudo-direct estimates and calculate asymptotic 95\% confidence intervals as $\hat{y}_{rc} \pm z_{0.975}\sqrt{\hat{V}_{r,cc}}$ with $\hat{V}_{r,cc}$ the diagonal entries of the sampled covariance matrix. 

We will also fit six different smoothing models to the pseudo-direct estimates to estimate the latent means in each region with corresponding 95\% credible intervals. They will be:

1. Univariate IID
2. Univariate BYM
3. Bivariate nonshared IID
4. Bivariate nonshared BYM
5. Bivariate shared IID
6. Bivariate shared BYM

All data will be the same size as the Kenya 2014 DHS (47 regions). The DGM parameters will be set equal to those estimated from models fit to the Kenya 2014 DHS HAZ/WAZ data. We run 100 simulations.

\newpage

# Results

## All results
```{r, results = "asis"}
# set dir to files dir
setwd(dropbox_dir)

# file list
fs <- grep("results", list.files(), value = TRUE)

## extract run number to name saved results
run_numbers <- c()
for (i in 1:length(fs)) {
    run_numbers <- c(run_numbers,
                     as.numeric(regmatches(fs[i], regexec("results_run-\\s*(.*?)\\s*\\.Rdata", fs[i]))[[1]][2]))
}

run_numbers <- sort(run_numbers)

# read in results and display in tables
tmps <- vector(mode = "list", length = length(run_numbers))
# which_runs <- which(run_numbers %in% c(2, 14))
which_runs <- 1:length(run_numbers)
for (i in which_runs) {
    # cat(paste(i, "\n"))
    run_number <- run_numbers[i]
    myfile <- paste0("results_run-",run_number,".Rdata")
    load(myfile)
    tmps[[i]] <- results_comp
    rm(results_comp)
    
    dgmnum <- res_run$dgm_number[res_run$run_number == run_number]
    
    run_results <- cbind(model = c(model_names[1], ""), tmps[[i]][[1]])
    if (length(tmps[[i]]) > 1) {
        for (j in 2:length(tmps[[i]])) {
            # extract latent mean results
            tmp_res <- cbind(model = c(model_names[j], ""), 
                             tmps[[i]][[j]][tmps[[i]][[j]]$param %in% c("HAZ latent means", "WAZ latent means"),])
            run_results <- rbind(run_results, tmp_res)
        }   
    }
    
    print(kable(tmps[[i]][[j]], format = "markdown", caption = paste0("Model: ", model_names[j]), digits = 3,
              col.names = c("parameter", "bias", "abs bias", "relative abs bias", "80% coverage", "80% width", 
                            "95% coverage", "95% width")))
}
```


<!-- ## Scenario 1 -->

<!-- ### Specification -->

<!-- This scenario will generate data from a DGM without a shared component, using parameters from a model fit to the Kenya 2014 DHS without a shared component. We will fit a model without a shared component. Thus, it is a correctly specified model for the DGM. -->

<!-- Here's a table with the parameter values for the DGM -->

<!-- ```{r} -->

<!-- # load parameter values -->
<!-- tmp <- read_rds(paste0("~/Dropbox/dissertation_2/survey-csmf/results/ken2014-hazwaz/", -->
<!--                        dgm_info$V_corr_lower[dgm_info$dgm_number == res_run$dgm_number[res_run$run_number == 1]], -->
<!--                        ".rds")) -->
<!-- params_to_extract <- c("beta", "sigma", "rho") -->
<!-- tmp_summary <- summary(tmp$mod[[1]], pars = params_to_extract, probs = c(0.5))$summary[, "50%"] -->
<!-- tmp_summary <- c(tmp_summary, lambda = 0) -->
<!-- kable(tmp_summary, format = "markdown", digits = 3) -->
<!-- ``` -->

<!-- We will fit this model to the data: -->

<!-- \eq{ -->
<!--     \mu^{HAZ}_r &= \alpha_1 + \sigma_1(\sqrt{1 - \rho_1}v_{1r}^* + \sqrt{\rho_1}u_{1r}^*) \\ -->
<!--     \mu^{WAZ}_r &= \alpha_2 + \sigma_2(\sqrt{1 - \rho_2}v_{2r}^* + \sqrt{\rho_2}u_{2r}^*) \\ -->
<!--     v_{1r} &\overset{iid}{\sim} N(0, 1) \\ -->
<!--     v_{2r} &\overset{iid}{\sim} N(0, 1) \\ -->
<!--     \bm{u}_1 &\sim ICAR(1) \\ -->
<!--     \bm{u}_2 &\sim ICAR(1)  -->
<!-- } -->

<!-- We will put flat priors on the $\alpha_c$, $\mbox{Beta}(1, 1)$ priors on the $\rho_c$, and half-Normal(0, 1) priors on the $\sigma_c$. -->

<!-- ### Results -->

<!-- ```{r} -->
<!-- # table of results -->
<!-- load("~/Dropbox/dissertation_2/survey-csmf/results/stage-2-simulation/results-diags_run-1.Rdata") -->
<!-- kable(results_comp, format = "markdown", digits = 3, -->
<!--       col.names = c("param", "mean abs bias", "mean rel bias", "mean cov 80%", "mean width 80%", "mean cov 95%", "mean width 95%")) -->
<!-- ``` -->

<!-- \newpage -->

<!-- ## Scenario 2 -->

<!-- ### Specification -->

<!-- This scenario will generate data from a DGM with a shared component, using parameters from a model fit to the Kenya 2014 DHS with a shared component. But we will fit a model without a shared component, thus it will be misspecified. -->

<!-- Here's a table with the parameter values for the DGM -->

<!-- ```{r} -->

<!-- # load parameter values -->
<!-- tmp <- read_rds(paste0("~/Dropbox/dissertation_2/survey-csmf/results/ken2014-hazwaz/", -->
<!--                        dgm_info$V_corr_lower[dgm_info$dgm_number == res_run$dgm_number[res_run$run_number == 2]], -->
<!--                        ".rds")) -->
<!-- params_to_extract <- c("beta", "sigma", "rho", "lambda") -->
<!-- tmp_summary <- summary(tmp$mod[[1]], pars = params_to_extract, probs = c(0.5))$summary[, "50%"] -->
<!-- kable(tmp_summary, format = "markdown", digits = 3) -->
<!-- ``` -->

<!-- We will fit this model to the data: -->

<!-- \eq{ -->
<!--     \mu^{HAZ}_r &= \alpha_1 + \sigma_1(\sqrt{1 - \rho_1}v_{1r}^* + \sqrt{\rho_1}u_{1r}^*) \\ -->
<!--     \mu^{WAZ}_r &= \alpha_2 + \sigma_2(\sqrt{1 - \rho_2}v_{2r}^* + \sqrt{\rho_2}u_{2r}^*) \\ -->
<!--     v_{1r} &\overset{iid}{\sim} N(0, 1) \\ -->
<!--     v_{2r} &\overset{iid}{\sim} N(0, 1) \\ -->
<!--     \bm{u}_1 &\sim ICAR(1) \\ -->
<!--     \bm{u}_2 &\sim ICAR(1)  -->
<!-- } -->

<!-- We will put flat priors on the $\alpha_c$, $\mbox{Beta}(1, 1)$ priors on the $\rho_c$, and half-Normal(0, 1) priors on the $\sigma_c$. -->

<!-- ### Results -->

<!-- ```{r} -->
<!-- rm(results_comp) -->
<!-- rm(diags_comp) -->

<!-- # table of results -->
<!-- load("~/Dropbox/dissertation_2/survey-csmf/results/stage-2-simulation/results-diags_run-1.Rdata") -->
<!-- kable(results_comp, format = "markdown", digits = 3, -->
<!--       col.names = c("param", "mean abs bias", "mean rel bias", "mean cov 80%", "mean width 80%", "mean cov 95%", "mean width 95%")) -->
<!-- ``` -->

<!-- \newpage -->

<!-- ## Scenario 3 -->

<!-- ## Scenario 4 -->

<!-- ## Scenario 5 -->

<!-- ## Scenario 6 -->

<!-- ## Scenario 7 -->

<!-- ## Scenario 8 -->

